---
title: "S0_filteringdata"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
```
# loading data
```{r}
theme_set(theme_classic())
df_genotype_tidy <- readRDS("C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/Data/thymus_genotypes.tidy.rds")
df_sample_info   <- readRDS("C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/Data/thymus_sample_info.rds")
df_snp_info      <- readRDS("C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/Data/thymus_snps_info.rds")
```

# Informatıon about Datasets:

df_genotype_tidy:
includes sample id, snps and their genotypes

df_sample_info: 
sample id, plant id, species, site_id, tag_site, lat_deg, lat_deci, log_deg, log_deci, altitude, chemotype

df_snp_info:
snps, contigs, position, reference, alternative

# Identify bad samples
```{r}
bad_samples <- c("tha151", "tha063","tha238", "tha193") #from replicability
dim(df_genotype_tidy)
#sort(unique(df_genotype_tidy$sample_id)) # includes replicants, t.vulgaris, t.embergeri and 70km away
```

# remove bad samples and replicants
```{r}
df_genotype_tidy_without_bad_replicants <- df_genotype_tidy %>% 
  filter(!(sample_id %in% bad_samples )) %>% 
  filter(!grepl('b', sample_id))
dim(df_genotype_tidy_without_bad_replicants)
head(df_genotype_tidy_without_bad_replicants)
#sort(unique(df_genotype_tidy_without_bad_replicants$sample_id))#258 samples(t.vul, t.embergeri, 70km away)
```

# Define samples only with known chemotypes (excl. 70km away and T. embergeri)
253 - 259 70km away
260 - 262 T. embergeri
```{r}
all_ids <- sort(unique(df_genotype_tidy_without_bad_replicants$sample_id))
length(all_ids)
#all_ids
sml_ids <- all_ids[1:248]
t_embergeri_ids <- all_ids[256:258]
km_aways_ids <- all_ids[249:255]

sml_ids # thymus vulgaris with known chemotypes
t_embergeri_ids # t.embergeri ids
km_aways_ids # t.vulgaris located 70km away without knowing chemotypes
```

# Remove 70km away samples and T.embergeri samples from data and calculating rate of known snps
```{r}
pd <- df_genotype_tidy_without_bad_replicants %>%
  filter(sample_id %in% sml_ids) %>%
  group_by(snp) %>%
  summarise(good = sum(!is.na(genotype)),
            rate = good / n(),
            nb2 = sum(ifelse(is.na(genotype),0, genotype =="2")),
            nb1 = sum(ifelse(is.na(genotype),0, genotype =="1")),
            nb0 = sum(ifelse(is.na(genotype),0, genotype =="0")),
            ) %>%
  left_join(df_snp_info) %>%
  {.}

head(pd)
```

```{r}
# for t.embergeri ids
pd_t.embergeri <- df_genotype_tidy_without_bad_replicants %>%
  filter(sample_id %in% t_embergeri_ids) %>%
  group_by(snp) %>%
  summarise(good = sum(!is.na(genotype)),
            rate = good / n(),
            nb2 = sum(ifelse(is.na(genotype),0, genotype =="2")),
            nb1 = sum(ifelse(is.na(genotype),0, genotype =="1")),
            nb0 = sum(ifelse(is.na(genotype),0, genotype =="0")),
            ) %>%
  left_join(df_snp_info) %>%
  {.}

head(pd_t.embergeri)
```

# calculating freq of reference and alternative allele
```{r}
# for t.vulgaris
pd <- pd %>%
  mutate(freq_ref=(2*nb0 + nb1)/(2*good),
         freq_alt=(2*nb2 + nb1)/(2*good))

# for t.embergeri
pd_t.embergeri <- pd_t.embergeri %>%
  mutate(freq_ref=(2*nb0 + nb1)/(2*good),
         freq_alt=(2*nb2 + nb1)/(2*good))

head(pd)
head(pd_t.embergeri)
```

```{r}
# plot for t.vulgaris
ggplot(pd, aes(rate)) + geom_histogram(binwidth = 0.01) + 
  geom_vline(xintercept = c(0.5, 0.6, 0.75, 0.8, 0.9), col = c('blue', 'red', "orange", "green", "purple"))

# plot for t.embergeri
ggplot(pd_t.embergeri, aes(rate)) + geom_histogram(binwidth = 0.01) +
  geom_vline(xintercept = c(0.5, 0.6, 0.75, 0.8, 0.9), col = c('blue', 'red', "orange", "green", "purple"))
```

# filtered snps data for t.vulgaris and t.embergeri with 0.5 cutoff
```{r}
# t.vulgaris
df_snp_filtered.t.vulgaris.5 <- pd %>% filter(rate > 0.5)

# t.embergeri
df_snp_filtered.t.embergeri.5 <- pd_t.embergeri %>% filter(rate > 0.5)

# number of snps for each species
dim(df_snp_filtered.t.vulgaris.5)
dim(df_snp_filtered.t.embergeri.5)
```

# save filtered snp files
```{r}
# cutoff .5
#write_rds(df_snp_filtered.t.vulgaris.5, "C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_snp_filtered.t.vulgaris.5.rds")
#write_rds(df_snp_filtered.t.embergeri.5, "C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_snp_filtered.t.embergeri.5.rds")
```

# Number of genotypes after cutoff
```{r}
# t.vulgaris
df_genotype_filtered.t.vulgaris.5 <- df_genotype_tidy_without_bad_replicants %>%
  filter(snp %in% df_snp_filtered.t.vulgaris.5$snp) %>%
  filter(sample_id %in% sml_ids)

# t.embergeri
df_genotype_filtered.t.embergeri.5 <- df_genotype_tidy_without_bad_replicants %>%
  filter(snp %in% df_snp_filtered.t.embergeri.5$snp) %>%
  filter(sample_id %in% t_embergeri_ids)

# number of genotypes per species
dim(df_genotype_filtered.t.vulgaris.5)
dim(df_genotype_filtered.t.embergeri.5)
```

# Save filtered genotypes files for 0.5 cutoff
```{r}
#write_rds(df_genotype_filtered.t.vulgaris.5, "C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_genotype_filtered.t.vulgaris.5")
#write_rds(df_genotype_filtered.t.embergeri.5, "C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_genotype_filtered.t.embergeri.5")
```