---
title: "S1_finding_polymorphisms"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
```

# loading data
```{r}
theme_set(theme_classic())

#☺ filtered snp data
df_snp_filtered.t.vulgaris.5 <- readRDS("C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_snp_filtered.t.vulgaris.5.rds")
df_snp_filtered.t.embergeri.5 <- readRDS("C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_snp_filtered.t.embergeri.5.rds")

# filtered genotype data
df_genotype_filtered.t.vulgaris.5 <- read_rds("C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_genotype_filtered.t.vulgaris.5")
df_genotype_filtered.t.embergeri.5   <- read_rds("C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_genotype_filtered.t.embergeri.5")

```

# FOR SHARED POLYMORPHISM

If I remove snps that dont have heterozygotes and homozygotes alternative allele

```{r}
# polymorphic sites in T.vulgaris
df_poly_t.vulgaris.5 <- df_snp_filtered.t.vulgaris.5 %>% filter((nb0 != 0 & nb1 != 0) | (nb1 != 0 & nb2 != 0))

# polymorphic sites in T.embergeri
df_poly_t.embergeri.5 <- df_snp_filtered.t.embergeri.5 %>% filter((nb0 != 0 & nb1 != 0) | (nb1 != 0 & nb2 != 0)) 

head(df_poly_t.vulgaris.5)
head(df_poly_t.embergeri.5)

# number of polymorphic sites
dim(df_poly_t.vulgaris.5)
dim(df_poly_t.embergeri.5)

# ratio of polymorphisms 
# for t.vulgaris
dim(df_poly_t.vulgaris.5)[1]/dim(df_snp_filtered.t.vulgaris.5)[1] # ~0.75
# for t.embergeri
dim(df_poly_t.embergeri.5)[1]/dim(df_snp_filtered.t.embergeri.5)[1] # ~0.23
```

# shared polymorphisms data
```{r}
# cutoff 0.5
df_shared_poly.5 <- merge(df_poly_t.vulgaris.5, df_poly_t.embergeri.5,by = "snp",suffixes = c(".t.vulgaris", ".t.embergeri"))

dim(df_shared_poly.5) # 4486 shared poly.
```

```{r}
# for shared polymorphisms, we dont know yet which one is derived. 

# alt freq distributions
df_shared_poly.5 %>% ggplot(aes(freq_alt.t.vulgaris)) + geom_histogram(bins = 50)

df_shared_poly.5 %>% ggplot(aes(freq_alt.t.embergeri)) + geom_histogram()
```

# poly. only found in one species
```{r}
# poly.only found in t.vulgaris
df_only_poly_t.vulgaris.5 <- df_poly_t.vulgaris.5 %>% filter(!(snp %in% df_shared_poly.5$snp))

# poly. only found in t.embergeri
df_only_poly_t.embergeri.5 <- df_poly_t.embergeri.5 %>% filter(!(snp %in% df_shared_poly.5$snp))

dim(df_only_poly_t.vulgaris.5)
dim(df_only_poly_t.embergeri.5)
```

# combine polymorphisms into one dataset.
```{r}
# combine two only poly data
df_merged_polymorphisms.5 <- full_join(df_poly_t.vulgaris.5, df_poly_t.embergeri.5[c(1,2,3,4,5,6,11,12)], by = "snp", suffix = c(".t.vulgaris", ".t.embergeri"))

# unknown snps for each species
df_merged_polymorphisms.5 %>%  filter(is.na(df_merged_polymorphisms.5$good.t.vulgaris))
df_merged_polymorphisms.5 %>%  filter(is.na(df_merged_polymorphisms.5$good.t.embergeri))
dim(df_merged_polymorphisms.5) # 35603 (only t.vulgaris) + 4486 (common) + 7855 (t.embergeri) CHECK!

# Adding additional information to data (about shared or only)
df_merged_polymorphisms.5 <- df_merged_polymorphisms.5 %>% mutate(poly_INFO = ifelse(df_merged_polymorphisms.5$snp %in% df_shared_poly.5$snp, "shared", ifelse(df_merged_polymorphisms.5$snp %in% df_genotype_filtered.t.vulgaris.5$snp, "t.vulgaris", "t.embergeri")))

head(df_merged_polymorphisms.5)
```

# ancestral - derived alleles
compare only polymorphic data for t.vulgaris with t.embergeri genotypes
```{r}
# cutoff 0.5
df_genotypes_t.embergeri.anc_der.5 <- df_genotype_filtered.t.embergeri.5 %>% 
  filter(snp %in% df_only_poly_t.vulgaris.5$snp)

genotype_matrix.t.embergeri_anc_der.5 <- spread(df_genotypes_t.embergeri.anc_der.5, key = snp, value = genotype)
#head(genotype_matrix.t.embergeri_anc_der.5[,1:5])

dim(genotype_matrix.t.embergeri_anc_der.5)  # we dont have information for 98 snps (which is unique for t.vulgaris)
dim(df_only_poly_t.vulgaris.5)
```

# Finding homozygous sites in T.embergeri
I put symbol as comment because the running time of the chunk is long.
```{r}
## find homozoygote reference and alternative sites in t.embergeri
#only_2s.5 <- genotype_matrix.t.embergeri_anc_der.5[,2:ncol(genotype_matrix.t.embergeri_anc_der.5)] %>% select_if(function(x) any(x != 0))
#head(only_2s.5) # there are 2248 sites
#only_0s.5 <- setdiff(names(genotype_matrix.t.embergeri_anc_der.5[,2:ncol(genotype_matrix.t.embergeri_anc_der.5)]), names(only_2s.5))

## Check the dimensions of the data for correctness
#length(only_0s.5)
#dim(only_2s.5)[2]
#dim(genotype_matrix.t.embergeri_anc_der.5)
#length(only_0s.5) + dim(only_2s.5)[2] + 1 == dim(genotype_matrix.t.embergeri_anc_der.5)[2] # check!

# there are 2248 homozygote alternative in t.embergeri
# there are 33256 homozygote reference
```

# Save files
```{r}
## homozygotes sites of the t.embergeri
#write_rds(only_2s.5, "only_2s_geno_of_T.embergeri.5.rds") 
#write_rds(only_0s.5, "only_0s_geno_of_T.embergeri.5.rds")
```

# Load the homozygous geno of T.embergeri
```{r}
only_2s.5 <- readRDS("C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/only_2s_geno_of_T.embergeri.5.rds")
only_0s.5 <- readRDS("C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/only_0s_geno_of_T.embergeri.5.rds")
```

# Adding ancestral information to only.poly data of T.vulgaris
```{r}
df_only_poly_t.vulgaris.5 <-df_only_poly_t.vulgaris.5 %>% 
  mutate(ancestral_derived = ifelse(snp %in% names(only_2s.5), "Ref is derived/alt is ancestral", ifelse(snp %in% only_0s.5, "Ref is ancestral/alt is derived", "Unknown"))) 

df_only_poly_t.vulgaris.5 %>% count(ancestral_derived == "Unknown")
df_only_poly_t.vulgaris.5 %>% count(ancestral_derived == "Ref is derived/alt is ancestral")
df_only_poly_t.vulgaris.5 %>% count(ancestral_derived == "Ref is ancestral/alt is derived")

dim(df_only_poly_t.vulgaris.5)[1] == 99 + 2248 + 33256 #check!

# 99 unknown, 2248 ref is derived, 33256 ref is ancestral

```

# adding ancestral/derived information to total polymorphism data which contains information about both t.vulgaris and t.embergeri
```{r}
df_merged_polymorphisms.5 <- df_merged_polymorphisms.5 %>% mutate(ancestral_derived = ifelse(snp %in% df_only_poly_t.vulgaris.5$snp, df_only_poly_t.vulgaris.5$ancestral_derived, "-"))

dim(df_merged_polymorphisms.5 %>% filter(ancestral_derived == "-"))[1] == 4486 + 7855 # check!
```

# ancestral freq vs. derived frequency into the merged_polymorphism and only_poly.t.vulgaris data
```{r}
df_merged_polymorphisms.5 <- df_merged_polymorphisms.5 %>% 
  mutate(ancestral_freq.t.vulgaris = ifelse(ancestral_derived == "Ref is ancestral/alt is derived", freq_ref.t.vulgaris, ifelse(ancestral_derived == "Ref is derived/alt is ancestral", freq_alt.t.vulgaris, freq_ref.t.vulgaris))) %>% 
  mutate(derived_freq.t.vulgaris = ifelse(ancestral_derived == "Ref is ancestral/alt is derived", freq_alt.t.vulgaris, ifelse(ancestral_derived == "Ref is derived/alt is ancestral", freq_ref.t.vulgaris, freq_alt.t.vulgaris)))

df_only_poly_t.vulgaris.5 <- df_only_poly_t.vulgaris.5 %>% 
  mutate(ancestral_freq = ifelse(ancestral_derived == "Ref is ancestral/alt is derived", freq_ref, ifelse(ancestral_derived == "Ref is derived/alt is ancestral", freq_alt, freq_ref))) %>% 
  mutate(derived_freq = ifelse(ancestral_derived == "Ref is ancestral/alt is derived", freq_alt, ifelse(ancestral_derived == "Ref is derived/alt is ancestral", freq_ref, freq_alt)))

head(df_merged_polymorphisms.5)
```

# Save created data
```{r}
# only poly.
#write_rds(df_only_poly_t.vulgaris.5, "C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_only_poly_t.vulgaris.5.rds")
#write_rds(df_only_poly_t.embergeri.5, "C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_only_poly_t.embergeri.5.rds")

# merged poly.
#write_rds(df_merged_polymorphisms.5, "C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_merged_polymorphisms.5.rds")

# save shared polymorphism files
#write_rds(df_shared_poly.5, "C:/Users/Aysevil/Dropbox/Aarhus Lectures/Project/tidy files/df_shared_poly.5.rds")
```











